---
title: "Prediction Assignment"
author: "Zhiwei Xu"
date: "9/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(caret)
```

## Prediction Assignment

Steps:

1. Read in train, test dataset
2. Exploratory analysis
3. Preprocessing: 
  + Convert classe to factor variable
  + Remove those columns that are NA in test set 
  + Remove timestamps, window indicator, X, and username as they are not informative predictors to be included in the model
4. Modeling:
  + rpart with CV
    + Best model: cp = 0.0356 
    + Model accuracy: 0.5066
    + Predict result: C A C A A C C A A A C C C A C A A A A C
  + Random Forest 
    + Best model: mtry = 27
    + Model accuracy: 0.9927 
    + Predict result: B A B A A E D B A A B C B A E E A B B B
  + GBM
    + Best model: shrinkage = 0.1, n.trees = 150, interaction.depth = 3, n.minobsinnode = 10
    + Model accuracy: 0.9598
    + Predict result: B A B A A E D B A A B C B A E E A B B B
5. Result:
  Confirm the predict results from RF and GBM by inputing them into the Course Project Prediction Quiz and get 100% accuracy rate
  
```{r}
##Import train set and test set
train <- read.csv("C:\\Users\\xuzhi\\Downloads\\pml-training.csv")
test <- read.csv("C:\\Users\\xuzhi\\Downloads\\pml-testing.csv")

##Preprocessing
#Convert a few vars to factors
train$new_window <- as.factor(train$new_window)
train$user_name <- as.factor(train$user_name)
train$classe <- as.factor(train$classe)

##Exploratory analysis
#summary(train)
#str(train)
  #num_window (1~864)
  #cvtd_timestamp 28/11/2011 14:15 ~ 	02/12/2011 13:35
  # adelmo carlitos  charles   eurico   jeremy    pedro 
  #   3892     3112     3536     3070     3402     2610

#Select the NA columns in test, these columns should not be included in the predictors 
na_count <-sapply(test, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
na_columns <- rownames(na_count%>%filter(na_count==20))
train <- train[, -which(names(train) %in% na_columns)]
#remove timestamp, window indicator, X, username
train <- train[, -which(names(train) %in% c("raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window","num_window","X","user_name"))]
  
##Covariant creation
#Don't see the need

##Modeling
#Regression
#Tree
train_control<- trainControl(method="cv", number=10)
rpart_mod <- train(classe ~ ., method="rpart", trControl=train_control, data=train)
plot(rpart_mod$finalModel)
text(rpart_mod$finalModel)
pred_rpart <- predict(rpart_mod,test)
#Bagging
#RF
rf_mod <- train(classe ~ ., method="rf", data=train)
pred_rf <- predict(rf_mod,test)
#Boosting
gbm_mod <- train(classe ~ ., method="gbm", data=train)
pred_gbm <- predict(gbm_mod,test)

#
pred_rpart
pred_rf
pred_gbm
```
